---
title: "Data Science in Context - Water"
author: "Julia Ferris"
date: "2023-10-02"
output:
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this code:

In my Data Science in Context presentation, I only used some of the graphs produced here. I also only included graphs with complete sets of data. Any missing values were not used. In the simplified data sets, E stands for estimated data, and I stands for imputed data.

## Data

```{r cars}
library(readr)
library(dplyr)
library(ggplot2)
fullWaterData <- read_csv("https://raw.githubusercontent.com/juliaDataScience-22/cuny-fall-23/manage-acquire-data/Water-Data-Sc-Context.csv")
fullWaterData$Country <- replace(fullWaterData$Country, fullWaterData$Country == "C\xf4te d'Ivoire", "Cote d'Ivoire")
fullWaterData$Country <- replace(fullWaterData$Country, fullWaterData$Country == "T\xfcrkiye", "Turkiye")
View(fullWaterData)

simplifiedWaterData <- data.frame(Country = fullWaterData$Country,
                          Year = fullWaterData$Year,
                          `Agricultural Water Withdrawal` = coalesce
                          (fullWaterData$`E - Agricultural water withdrawal as % of total renewable water resources`, 
                            fullWaterData$`I - Agricultural water withdrawal as % of total renewable water resources`),
                          `Industrial Water Use` = coalesce(fullWaterData$`E - SDG 6.4.1. Industrial Water Use Efficiency`, 
                                                            fullWaterData$`I - SDG 6.4.1. Industrial Water Use Efficiency`),
                          `Irrigated Agriculture Water Use` = coalesce
                          (fullWaterData$`E - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`, 
                            fullWaterData$`I - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`),
                          `Services Water Use` = fullWaterData$`E - SDG 6.4.1. Services Water Use Efficiency`,
                          `Water Use` = coalesce(fullWaterData$`E - SDG 6.4.1. Water Use Efficiency`, 
                                                 fullWaterData$`I - SDG 6.4.1. Water Use Efficiency`),
                          `Water Stress` = coalesce(fullWaterData$`E - SDG 6.4.2. Water Stress`, 
                                                    fullWaterData$`I - SDG 6.4.2. Water Stress`))
View(simplifiedWaterData)

usWaterData <- (fullWaterData[1751:1760,])
View(usWaterData)

simplifiedUSWaterData <- simplifiedWaterData[1751:1760,]
View(simplifiedUSWaterData)

canadaWaterData <- simplifiedWaterData[301:310,]
View(canadaWaterData)

mexicoWaterData <- simplifiedWaterData[1071:1080,]
View(mexicoWaterData)
```

## Visualize Trends


```{r pressure, fig.show="hold", out.width="50%"}
ggplot(fullWaterData, aes(Year, `E - Agricultural water withdrawal as % of total renewable water resources`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,3)], desc(`E - Agricultural water withdrawal as % of total renewable water resources`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `I - Agricultural water withdrawal as % of total renewable water resources`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,4)], desc(`I - Agricultural water withdrawal as % of total renewable water resources`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `E - SDG 6.4.1. Industrial Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,5)], desc(`E - SDG 6.4.1. Industrial Water Use Efficiency`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `I - SDG 6.4.1. Industrial Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,6)], desc(`I - SDG 6.4.1. Industrial Water Use Efficiency`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `E - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,7)], desc(`E - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `I - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,8)], desc(`I - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `E - SDG 6.4.1. Services Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,9)], desc(`E - SDG 6.4.1. Services Water Use Efficiency`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `E - SDG 6.4.1. Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,10)], desc(`E - SDG 6.4.1. Water Use Efficiency`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `I - SDG 6.4.1. Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,11)], desc(`I - SDG 6.4.1. Water Use Efficiency`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `E - SDG 6.4.2. Water Stress` , color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,12)], desc(`E - SDG 6.4.2. Water Stress`)) |> head(n = 30)

ggplot(fullWaterData, aes(Year, `I - SDG 6.4.2. Water Stress` , color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,13)], desc(`I - SDG 6.4.2. Water Stress`)) |> head(n = 30)

```


## Time Series Analysis

```{r simplified-graphs, fig.show="hold", out.width="50%"}
ggplot(simplifiedWaterData, aes(Year, `Agricultural.Water.Withdrawal`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(simplifiedWaterData[,c(1,2,3)], desc(`Agricultural.Water.Withdrawal`)) |> head(n = 30)

ggplot(simplifiedWaterData, aes(Year, Industrial.Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(simplifiedWaterData[,c(1,2,4)], desc(Industrial.Water.Use)) |> head(n = 30)

ggplot(simplifiedWaterData, aes(Year, Irrigated.Agriculture.Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(simplifiedWaterData[,c(1,2,5)], desc(Irrigated.Agriculture.Water.Use)) |> head(n = 30)

ggplot(simplifiedWaterData, aes(Year, Services.Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(simplifiedWaterData[,c(1,2,6)], desc(Services.Water.Use)) |> head(n = 30)

ggplot(simplifiedWaterData, aes(Year, Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(simplifiedWaterData[,c(1,2,7)], desc(Water.Use)) |> head(n = 30)

ggplot(simplifiedWaterData, aes(Year, Water.Stress, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(simplifiedWaterData[,c(1,2,8)], desc(Water.Stress)) |> head(n = 30)

```
## US Water Stress
```{r water-us, fig.show="hold", out.width="50%"}
ggplot(usWaterData, aes(Year, `E - Agricultural water withdrawal as % of total renewable water resources`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(usWaterData[,c(1,2,3)], desc(`E - Agricultural water withdrawal as % of total renewable water resources`)) |> head(n = 30)

ggplot(usWaterData, aes(Year, `I - Agricultural water withdrawal as % of total renewable water resources`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(usWaterData[,c(1,2,4)], desc(`I - Agricultural water withdrawal as % of total renewable water resources`)) |> head(n = 30)

ggplot(usWaterData, aes(Year, `E - SDG 6.4.1. Industrial Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(usWaterData[,c(1,2,5)], desc(`E - SDG 6.4.1. Industrial Water Use Efficiency`)) |> head(n = 30)

ggplot(usWaterData, aes(Year, `E - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(usWaterData[,c(1,2,7)], desc(`E - SDG 6.4.1. Irrigated Agriculture Water Use Efficiency`)) |> head(n = 30)

ggplot(usWaterData, aes(Year, `E - SDG 6.4.1. Services Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,9)], desc(`E - SDG 6.4.1. Services Water Use Efficiency`)) |> head(n = 30)

ggplot(usWaterData, aes(Year, `I - SDG 6.4.1. Water Use Efficiency`, color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,11)], desc(`I - SDG 6.4.1. Water Use Efficiency`)) |> head(n = 30)

ggplot(usWaterData, aes(Year, `E - SDG 6.4.2. Water Stress` , color = Country)) +
  geom_point() + theme(legend.position = "none")
arrange(fullWaterData[,c(1,2,12)], desc(`E - SDG 6.4.2. Water Stress`)) |> head(n = 30)

```

## US Water Stress - Simplified

```{r more-water-data, fig.show="hold", out.width="50%"}
ggplot(simplifiedUSWaterData, aes(Year, `Agricultural.Water.Withdrawal`, color = Country)) +
  geom_point() + theme(legend.position = "none")

ggplot(simplifiedUSWaterData, aes(Year, Industrial.Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")

ggplot(simplifiedUSWaterData, aes(Year, Irrigated.Agriculture.Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")

ggplot(simplifiedUSWaterData, aes(Year, Services.Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")

ggplot(simplifiedUSWaterData, aes(Year, Water.Use, color = Country)) +
  geom_point() + theme(legend.position = "none")

ggplot(simplifiedUSWaterData, aes(Year, Water.Stress, color = Country)) +
  geom_point() + theme(legend.position = "none")

```

``` {r all-data, fig.show="hold", out.width="50%"}
ggplot() +
  geom_point(data = canadaWaterData, mapping = aes(Year, `Agricultural.Water.Withdrawal`, color = Country)) +
  geom_point(data = simplifiedUSWaterData, mapping = aes(Year, `Agricultural.Water.Withdrawal`, color = Country)) +
  geom_point(data = mexicoWaterData, mapping = aes(Year, `Agricultural.Water.Withdrawal`, color = Country))

ggplot() +
  geom_point(data = canadaWaterData, mapping = aes(Year, `Industrial.Water.Use`, color = Country)) +
  geom_point(data = simplifiedUSWaterData, mapping = aes(Year, `Industrial.Water.Use`, color = Country)) +
  geom_point(data = mexicoWaterData, mapping = aes(Year, `Industrial.Water.Use`, color = Country))

ggplot() +
  geom_point(data = canadaWaterData, mapping = aes(Year, `Irrigated.Agriculture.Water.Use`, color = Country)) +
  geom_point(data = simplifiedUSWaterData, mapping = aes(Year, `Irrigated.Agriculture.Water.Use`, color = Country)) +
  geom_point(data = mexicoWaterData, mapping = aes(Year, `Irrigated.Agriculture.Water.Use`, color = Country))

ggplot() +
  geom_point(data = canadaWaterData, mapping = aes(Year, `Services.Water.Use`, color = Country)) +
  geom_point(data = simplifiedUSWaterData, mapping = aes(Year, `Services.Water.Use`, color = Country)) +
  geom_point(data = mexicoWaterData, mapping = aes(Year, `Services.Water.Use`, color = Country))

ggplot() +
  geom_point(data = canadaWaterData, mapping = aes(Year, `Water.Use`, color = Country)) +
  geom_point(data = simplifiedUSWaterData, mapping = aes(Year, `Water.Use`, color = Country)) +
  geom_point(data = mexicoWaterData, mapping = aes(Year, `Water.Use`, color = Country))

ggplot() +
  geom_point(data = canadaWaterData, mapping = aes(Year, `Water.Stress`, color = Country)) +
  geom_point(data = simplifiedUSWaterData, mapping = aes(Year, `Water.Stress`, color = Country)) +
  geom_point(data = mexicoWaterData, mapping = aes(Year, `Water.Stress`, color = Country))
```

```{r comparison, fig.show="hold", out.width="33%"}
ggplot(canadaWaterData, aes(Year, `Agricultural.Water.Withdrawal`, color = Country)) +
  geom_point(color = "blue") + geom_smooth(color = "blue") +
  theme(legend.position = "none") +
  labs(title = "Canada")
ggplot(simplifiedUSWaterData, aes(Year, `Agricultural.Water.Withdrawal`, color = Country)) +
  geom_point(color = "red") + geom_smooth(color = "red") + 
  theme(legend.position = "none") +
  labs(title = "U.S.")
ggplot(mexicoWaterData, aes(Year, `Agricultural.Water.Withdrawal`, color = Country)) +
  geom_point(color = "green") + geom_smooth(color = "green") + 
  theme(legend.position = "none") +
  labs(title = "Mexico")

ggplot(canadaWaterData, aes(Year, Industrial.Water.Use, color = Country)) +
  geom_point(color = "blue") + geom_smooth(color = "blue") + 
  theme(legend.position = "none") +
  labs(title = "Canada")
ggplot(simplifiedUSWaterData, aes(Year, Industrial.Water.Use, color = Country)) +
  geom_point(color = "red") + geom_smooth(color = "red") +
  theme(legend.position = "none") +
  labs(title = "U.S.")
ggplot(mexicoWaterData, aes(Year, Industrial.Water.Use, color = Country)) +
  geom_point(color = "green") + geom_smooth(color = "green") +
  theme(legend.position = "none") +
  labs(title = "Mexico")

ggplot(canadaWaterData, aes(Year, Irrigated.Agriculture.Water.Use, color = Country)) +
  geom_point(color = "blue") + geom_smooth(color = "blue") +
  theme(legend.position = "none") +
  labs(title = "Canada")
ggplot(simplifiedUSWaterData, aes(Year, Irrigated.Agriculture.Water.Use, color = Country)) +
  geom_point(color = "red") + geom_smooth(color = "red") +
  theme(legend.position = "none") +
  labs(title = "U.S.")
ggplot(mexicoWaterData, aes(Year, Irrigated.Agriculture.Water.Use, color = Country)) +
  geom_point(color = "green") + geom_smooth(color = "green") +
  theme(legend.position = "none") +
  labs(title = "Mexico")

ggplot(canadaWaterData, aes(Year, Services.Water.Use, color = Country)) +
  geom_point(color = "blue") + geom_smooth(color = "blue") +
  theme(legend.position = "none") +
  labs(title = "Canada")
ggplot(simplifiedUSWaterData, aes(Year, Services.Water.Use, color = Country)) +
  geom_point(color = "red") + geom_smooth(color = "red") +
  theme(legend.position = "none") +
  labs(title = "U.S.")
ggplot(mexicoWaterData, aes(Year, Services.Water.Use, color = Country)) +
  geom_point(color = "green") + geom_smooth(color = "green") +
  theme(legend.position = "none") +
  labs(title = "Mexico")

ggplot(canadaWaterData, aes(Year, Water.Use, color = Country)) +
  geom_point(color = "blue") + geom_smooth(color = "blue") +
  theme(legend.position = "none") +
  labs(title = "Canada")
ggplot(simplifiedUSWaterData, aes(Year, Water.Use, color = Country)) +
  geom_point(color = "red") + geom_smooth(color = "red") +
  geom_smooth() + theme(legend.position = "none") +
  labs(title = "U.S.")
ggplot(mexicoWaterData, aes(Year, Water.Use, color = Country)) +
  geom_point(color = "green") + geom_smooth(color = "green") +
  theme(legend.position = "none") +
  labs(title = "Mexico")

ggplot(canadaWaterData, aes(Year, Water.Stress, color = Country)) +
  geom_point(color = "blue") + geom_smooth(color = "blue") +
  theme(legend.position = "none") +
  labs(title = "Canada")
ggplot(simplifiedUSWaterData, aes(Year, Water.Stress, color = Country)) +
  geom_point(color = "red") + geom_smooth(color = "red") +
  theme(legend.position = "none") +
  labs(title = "U.S.")
ggplot(mexicoWaterData, aes(Year, Water.Stress, color = Country)) +
  geom_point(color = "green") + geom_smooth(color = "green") +
  theme(legend.position = "none") +
  labs(title = "Mexico")

```
